Hierarchical Principal Manifold Estimation

### 2023-08-15 paper/main.tex
add preliminary hierarchical pme discussion

While this approach to tuning the model parameters is beneficial because it allows a complete comparison of all parameter values under consideration, the additive model described in equation \ref{eq:2} requires many more parameters, meaning that taking the same tuning approach would result in the need to find an optimal point in a very high-dimensional space. This would greatly increase the computational complexity of the algorithm, likely to the point of infeasibility in an practical situation. Instead, the proposed algorithm avoids this problem by taking a greedy approach to tuning the model, where the smoothing parameters for each model are chosen individually, rather than simultaneously. The proposed algorithm is given as Algorithm \ref{alg:hpme}.

\begin{algorithm}
\caption{Backfitting Algorithm for Additive Spline Model}\label{alg:hpme}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, group and individual indicator values $\left\{g(j)\right\}_{j=1}^{N}$ and $\left\{i(j)\right\}_{j=1}^{N}$, and $\epsilon$.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$.}

>>>>

While this approach to tuning the model parameters is beneficial because it allows a complete comparison of all parameter values under consideration, the additive model described in equation \ref{eq:2} requires many more parameters, meaning that taking the same tuning approach would result in the need to find an optimal point in a very high-dimensional space. This would greatly increase the computational complexity of the algorithm, likely to the point of infeasibility in an practical situation. Instead, the proposed algorithm avoids this problem by taking a greedy approach to tuning the model, where the smoothing parameters for each model are chosen individually, rather than simultaneously.

Thus, the hierarchical PME algorithm proposed here is composed of initialization and data reduction and fitting steps. The first step uses the PME algorithm to estimate an embedding map from one image. The HDMDE algorithm, detailed in \cite{mengPrincipalManifoldEstimation2021}, is used to reduce the data from each image into a set of component centers and associated weights. Then, the embedding map estimated using the PME algorithm can be used to provide initial parameterizations of the component centers for every image. These initial parameterizations are then used in the fitting step.

Like the PME algorithm, the fitting step uses an iterative approach to alternate between using the parameterizations to estimate an embedding function with the form of the additive spline model in equation \ref{eq:2}, then using this newly generated embedding function to reestimate the most appropriate parameterizations for the reduced set of components identified in the data reduction process. Once the error of the embedding function as measured by the sum of squared distances measured at the component centers declines by less than a given threshold $\epsilon$, the iterative procedure stops and the embedding function and associated parameterizations are finalized. The details of this process are given in Algorithm \ref{alg:hpme}.

\begin{algorithm}
\caption{Hierarchical Principal Manifold Estimation Algorithm}\label{alg:hpme}
  \KwData{Data points $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_j}$ in $\mathbb{R}^{D}$; a positive integer $N_0 < \sum_{j=1}^{N}M_j$; $\alpha, \epsilon, \epsilon^{*} \in (0, 1)$; candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}, \left\{\gamma_p\right\}_{p=1}^{P}$; $itr \geq 1$, which is the maximum number of iterations allowed.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$, .}
